Supplementary material: When does vapor pressure deficit drive or reduce evapotranspiration?

### 2019-09-04 doc/paper/agu-james-submission.tex
done with response to review 1

  \begin{equation}
    g_{1}  \propto \sqrt{\frac{3 \, \Gamma^* \,
        \lambda}{1.6}},\footnote{Note this expression has units of of
      (mmol mol$^{-1}$)$^{1/2}$, but this can be converted to
      Pa$^{1/2}$ using the ideal gas law.}
    \label{slope}
  \end{equation}
% \end{linenomath*}

where $\Gamma^*$ is the CO$_2$ compensation point for photosynthesis
(without dark respiration), and $\lambda$ is the marginal water cost
of leaf carbon
($\frac{\partial \; \text{transpiration}}{\partial \; CO_2 assimilation}$)
\cite{Farquhar_1980, Katul_2009}. So,
$g_{1}$ is a leaf-scale term reflecting the trade-off of water for
carbon uptake. The higher $g_{1}$, the more open the stomata and
the more they release water in exchange for carbon.

@@

  providing analytical framework for ecosystem response to
  atmospheric demand with environmental conditions held
  fixed. There are a few subtleties to taking the derivative in
  Eq. (\ref{d_et}): $\Delta$ ($\frac{d e_{s}}{d T}$) and $VPD$ are
  functionally related, so while taking the derivative we evaluate
  $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET}
  {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD}
  \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH}
  \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$
      fixed}}$. $RH$ and $e_s$ are assumed to be approximately
  independent, which is supported by data (not shown).

  This derivation relied either implicitly or explicitly on several
  assumptions. First, we assume that VPD at the leaf surface is the
  same as VPD at measurement height; physically this implies that
  leaves are perfectly coupled to the atmosphere. In reality, for some
  conditions and plant types the leaves can become decoupled from the
  boundary layer \cite{De_2017, Medlyn_2017}. Therefore, our
  derivation will be most applicable in times like the growing season
  (when we also expect uWUE to be most valid), when relatively high
  insolation induces instability and convective boundary layers, and
  we would expect the surface to be generally well coupled. An
  additional assumption in the formulation of uWUE \cite{Zhou_2014,
    Zhou_2015} and \citeA{Medlyn_2017}'s stomatal conductance model is
  that direct soil evaporation contributions to ET remain small
  relative to transpiration. Again, this should be more true during
  the growing season. The ratio of evaporation to transpiration may
  increase immediately after rainfall events due to high soil moisture
  and ponding, but VPD is generally low anyways during these
  times. However, some plant types allow for systematically larger
  contributions of evaporation in ET, particularly those with sparse
  canopies and smaller relative amounts of transpiration. We therefore
  might expect that the theory will be most applicable to forest PFTs,
  which will be most strongly coupled to the boundary layer due to
  larger surface roughness, and will also generally have the highest
  ratios of transpiration to evaporation. Finally, we assume that
  $g_1$ and uWUE are constant with respect to the conceptual VPD
  perturbation. Both quantities have been shown to be relatively
  constant with respect to changes in VPD \cite{Franks_2017,
    Zhou_2014}. These parameters will however vary with plant species
  and characteristics \cite<e.g. wood density, >[]{Lin_2015}, as well
  as environmental conditions including soil water content and
  temperature \cite{Lin_2015, Manzoni2013}. Exploring possible soil
  moisture (in)dependence of the plant parameters (g$_1$ and uWUE) is
  particularly interesting because soil moisture only enters the
  partial derivative directly through these plant parameters. If the
  plant parameters are weak functions of soil moisture then the theory
  can be directly applied to a broader range of conceptual VPD
  scenarios, including observed compound events between high VPD and
  low soil moisture \cite{Zhou_2019}. Supplementary material for this
  manuscript explores the soil moisture dependence of the plant
  parameters, but excessive noise and inconsistencies preclude
  conclusions about the nature of any dependence. We provide it in
  case it is of interest to the reader, and to motivate future
  research.

  Because Eq. (\ref{d_et}) is a partial derivative, it's utility is as
  a conceptual model for the change in ET in response to VPD with all
  other variables held fixed. This provides a useful tool for
  identifying the effect of VPD on ET through atmospheric demand and
  plant response, and allows a practitioner to disentangle complicated
  feedbacks when many quantities co-vary. However users of
  Eq. (\ref{d_et}) should take care that their interpretation matches
  the assumptions inherent in a partial derivative. For example,
  results will only be as valid as the assumption that g$_1$ and uWUE
  (and by extension $\lambda$) are fixed with respect to the user's
  conceptual change in VPD. Care must also be taken with possible
  indirect effects associated with a change in VPD: for example, a
  change in ET induced by a change in VPD can cause a change in
  surface temperature, which would drive a change in net
  radiation. These types of indirect effects and feedbacks are not
  considered in Eq. (\ref{d_et}): temperature (a variable) is
  mathematically fixed.

@@

\citeA{Medlyn_2017}'s ecosystem-scale results did not display the
robust relationship between plant type and and plant parameter (g$_1$)
that was demonstrated by \citeA{Zhou_2015} for uWUE. Given that there
is a robust relationship between g$_1$ and plant type in
\citeA{Lin_2015}'s leaf-scale results, and that \citeA{Zhou_2015}'s ecosystem-scale results
rely on the same underlying theory as \citeA{Medlyn_2017}'s results,
the ambiguous results in \citeA{Medlyn_2017} could be due to noise in
ecosystem scale observations, the consequences of imposing a model
structure in some of their g$_1$ estimations, or real ecosystem-scale
within-plant type variability (e.g. multiple species impacting the
relationship within the eddy-covariance footprint). These differences
between leaf-scale behavior and ecosystem scale behavior highlight the
importance of understanding how and why slope parameters show
increased variability at the ecosystem scale, as argued by
\citeA{Medlyn_2017}. However, if we assume that there is some real
analogy between leaf scale results in \citeA{Lin_2015} and ecosystem
scale behavior, and that some of the ambiguity in \citeA{Medlyn_2017}
is due to model and/or observational error, then we can use the
relationships between g$_1$ and plant characteristics defined by
\citeA{Lin_2015} to frame the ET response to VPD. This is subject to
the strong caveat that we still do not fully understand g$_1$ behavior
at the ecosystem scale. Extrapolating \citeA{Lin_2015}'s results
(e.g. Fig. 2) to the ecosystem scale would reveal the following
relationships between g$_1$ and plant types:

@@

So far we have discussed the role of different parameters in
modulating the ET response. However, the form of the ET response is
imposed by our choice of stomatal conductance model. We now examine
how stomatal model choice, and specifically the exponent of the VPD
dependence in the stomatal conductance model, can impact the general
form of the ET-VPD relationship. There is a theoretical basis for the
square root VPD dependence in both the stomatal conductance model and
uWUE based on the assumption that stomata behave to maximize carbon
gain while minimizing water loss, which observations also generally
support \cite{Lloyd_1991, MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015,
  Medlyn_2017}. However, some purely empirical results that fit the
exponent of the VPD dependence to data have shown that it may vary
slightly from 1/2, suggesting that stomata, as well as ecosystem-scale
quantities based on stomata theory, may not always function optimally
\cite{Zhou_2015, Lin_2018}. Specifically with regards to uWUE, one
would not expect that this ecosystem scale WUE quantity will respond
to VPD exactly analogously to stomata. Direct soil evaporation's
contributions to ET should shift the exponent of the VPD
dependence. \citeA{Zhou_2015}'s results corroborate this: they found a
mean empirically fit exponential VPD dependence of 0.55, varying
slightly from the theoretically optimal value of of 0.5 for AmeriFlux
sites. \citeA{Lin_2018}'s results also show variance in the empirical
exponent of the VPD dependence in the stomatal conductance model, but
interpretation of this variance is more difficult as \citeA{Lin_2018}
do not handle the GPP dependence of stomatal conductance in a directly
analogous manner to the optimal theory in \citeA{MEDLYN_2011} and
\citeA{Medlyn_2017}. Regardless, given that these recent results on
the relationship between VPD, GPP, and ET \cite{MEDLYN_2011,
  Zhou_2014, Zhou_2015, Medlyn_2017} form the backbone of our analysis
and are what allowed us to derive an explicit ET expression for the
first time (Eq. (\ref{et})), we will analyze if and how assumptions
about the exponent of the VPD dependence impact the shape of the
ET-VPD dependence. This analysis is also important to understand
whether the choice of stomatal conductance model alters the
fundamental behavior of the ET-VPD relationships, as many commonly
used models utilize a VPD exponent other than the 1/2 suggested by
optimal theory \cite<e.g. >[ which uses an exponent of
1]{Leuning_1990}. It is useful to understand how the choice of model
imposes the functional form of the ET-VPD relationship, given its
fundamental importance for land-atmosphere interactions.

@@

\begin{figure}
  \centering
  \centerline{\includegraphics[width=0.75\textwidth]{./concave.pdf}}
  \caption{ Solutions corresponding to inflection points between
    concave up and concave down ET-VPD curves (Eq. (\ref{curves}))
    for three specific scenarios. Solutions are defined in terms of a
    non-dimensional VPD ($VPD^m/g_*$), but to aide physical
    interpretation the horizontal axis is additionally provided in terms of dimensionalized VPD assuming $m=1/2$ and
    $g_*=110\; Pa^{1/2}$ (average of all PFT $g_1$). The vertical axis
    has a different interpretation depending on the solution
    curve. For the blue line ($m$ varying), it corresponds to $m$,
    for the orange line ($n$ varying) it corresponds to $n$, and for
    the green line it corresponds to the value of both $n$ and $m$
    ($n=m$). Regions of the parameter space that correspond to
    concave up and concave down results are labeled.}
  \label{concave}
\end{figure}

@@

While it is possible (and even likely) that in the future some other
form of VPD dependence is derived, at present \cite{MEDLYN_2011} and
\cite{Zhou_2014} established n=m=1/2 as the most likely candidate
given current theory and empirical data. Additionally, we argue that a
concave up result matches physical intuition more than a concave down
result. Plants must maintain nutrient and sugar transport through the
phloem and xylem. To accomplish this, stomata must remain slightly
open \cite{De_2013, Nikinmaa_2013, Ryan_2014}. Furthermore, even if
complete stomatal closure were possible, cuticular water loss and [at
the ecosystem-scale] direct soil evaporation are still sources of ET
which increase with VPD, independent of stomatal closure. Therefore,
in the limit as VPD becomes large and we assume plants are exercising
all strategies to reduce ET, any further increase in VPD should result
in an increase in ET through cuticular water loss and/or direct soil
evaporation. This inevitable transition from conditions when stomata
respond strongly to VPD to conditions when stomata response is
asymptoting towards full closure would cause a concave up ET-VPD
curve, which is matched by our theory. In short, plant response
becomes more limited as VPD increases, while atmospheric demand
monotonically increases with VPD, leading to the result that
atmospheric demand dominates plant response when atmospheric demand is
high.

This analysis allows us to understand the theoretical shape of the ET
response to VPD with environmental conditions held
fixed. Accomplishing this with purely statistical methods applied to
flux observations would be very difficult, given the relatively fast
time scale of plant response and the non-stationarity of [solar
forced] environmental conditions over the relatively coarse (half
hourly) flux estimates (which is required to obtain robust
eddy-covariance statistics). Our results on the shape of the ET-VPD
curve with environmental conditions held fixed can be built upon with
future work examining how changes in VPD and environmental conditions
(e.g. soil water storage) feedback upon one another. In the soil water
storage example, over very long time scales extremely high VPD
perturbations coupled with no precipitation could result in decreases
in soil water storage such that water becomes limiting. This could be
represented by an extension of our framework in which $\lambda$ is
allowed to decrease with decreasing soil water, increasing uWUE and
decreasing g$_1$. Here, we focus our results by framing them with
previous literature to build baseline intuition for ET-VPD dependence.

@@

\acknowledgments This work used eddy covariance data acquired and
shared by the FLUXNET community, including these networks: AmeriFlux,
AfriFlux, AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC,
OzFlux-TERN, TCOS-Siberia, and USCCC. The ERA-Interim reanalysis data
are provided by ECMWF and processed by LSCE. The FLUXNET eddy
covariance data processing and harmonization was carried out by the
European Fluxes Database Cluster, AmeriFlux Management Project, and
Fluxdata project of FLUXNET, with the support of CDIAC and ICOS
Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux
offices. We would like to thank Dr. Trevor Keenan for providing tools
for citation of the FLUXNET2015 dataset, available at
\url{https://github.com/trevorkeenan/FLUXNET_citations}. This material
is based upon work supported by the National Science Foundation
Graduate Research Fellowship under Grant No. DGE 16-44869.

>>>>

  \begin{equation}
    g_{1}  \propto \sqrt{\frac{3 \, \Gamma^* \,
        \lambda}{1.6}},
    \label{slope}
  \end{equation}
% \end{linenomath*}

  where g$_1$ is presented units of $\sqrt{\text{mol fraction}}$ for
  convenience (which can be converted to $\sqrt{Pa}$ using the ideal
  gas law and ambient pressure), $\Gamma^*$ is the CO$_2$ compensation
  point for photosynthesis (without dark respiration), and $\lambda$
  is the marginal water cost of leaf carbon
  ($\frac{\partial \; \text{transpiration}}{\partial \; CO_2
    assimilation}$) \cite{Farquhar_1980, Katul_2009}. So, $g_{1}$ is a
  leaf-scale term reflecting the trade-off of water for carbon
  uptake. The higher $g_{1}$, the more open the stomata and the more
  they release water in exchange for carbon.

@@

  providing analytical framework for ecosystem response to atmospheric
  demand with environmental conditions held fixed. There are a few
  subtleties to taking the derivative in Eq. (\ref{d_et}): $\Delta$
  ($\frac{d e_{s}}{d T}$) and $VPD$ are functionally related, so while
  taking the derivative we evaluate
  $\frac{\partial \; ET}{\partial \; VPD} = \frac{\partial \; ET}
  {\partial \; e_s} \frac{\partial \; e_s}{\partial \; VPD}
  \Big|_{\text{RH fixed}} + \frac{\partial \; ET}{\partial \; RH}
  \frac{\partial \; RH}{\partial \; VPD} \Big|_{\text{$e_s$
      fixed}}$. $RH$ and $e_s$ are assumed to be approximately
  independent, which is supported by data (see supplemental material).

  This derivation relied either implicitly or explicitly on several
  assumptions. First, we assume that VPD at the leaf surface is the
  same as VPD at measurement height; physically this implies that
  leaves are perfectly coupled to the atmosphere. In reality, for some
  conditions and plant types the leaves can become decoupled from the
  boundary layer \cite{De_2017, Medlyn_2017}. Therefore, our
  derivation will be most applicable in times like the growing season
  (when we also expect uWUE to be most valid), when relatively high
  insolation induces instability and convective boundary layers, and
  we would expect the surface to be generally well coupled. An
  additional assumption in the formulation of uWUE \cite{Zhou_2014,
    Zhou_2015} and the stomatal conductance model of
  \citeA{Medlyn_2017} is that direct soil evaporation contributions to
  ET remain small relative to transpiration. Again, this should be
  more true during the growing season. The ratio of evaporation to
  transpiration may increase immediately after rainfall events due to
  high soil moisture, ponding, and interception, but VPD is generally
  low anyways during these times. However, some plant types allow for
  systematically larger contributions of evaporation in ET,
  particularly those with sparse canopies and smaller relative amounts
  of transpiration. We therefore might expect that the theory will be
  most applicable to forest PFTs, which will be most strongly coupled
  to the boundary layer due to larger surface roughness, and will also
  generally have the highest ratios of transpiration to
  evaporation. Finally, we assume that $g_1$ and uWUE are constant
  with respect to the conceptual VPD perturbation. Both quantities
  have been shown to be relatively constant with respect to changes in
  VPD \cite{Franks_2017, Zhou_2014}. These parameters will however
  vary with plant species and characteristics \cite<e.g. wood density,
  >[]{Lin_2015}, as well as environmental conditions including soil
  water content and temperature \cite{Lin_2015,
    Manzoni2013}. Exploring possible soil moisture (in)dependence of
  the plant parameters (g$_1$ and uWUE) is particularly interesting
  because soil moisture only enters the partial derivative directly
  through these plant parameters. If the plant parameters are weak
  functions of soil moisture then the theory can be directly applied
  to a broader range of conceptual VPD scenarios, including observed
  compound events between high VPD and low soil moisture
  \cite{Zhou_2019}. Supplementary material for this manuscript
  explores the soil moisture dependence of the plant parameters, but
  excessive noise and inconsistencies preclude conclusions about the
  nature of any dependence. We provide it in case it is of interest to
  the reader, and to motivate future research.

  Because Eq. (\ref{d_et}) is a partial derivative, its utility is as
  a conceptual model for the change in ET in response to VPD with all
  other variables held fixed. This provides a useful tool for
  identifying the effect of VPD on ET through atmospheric demand and
  plant response, and allows a practitioner to disentangle complicated
  feedbacks when many quantities co-vary. However users of
  Eq. (\ref{d_et}) should take care that their interpretation matches
  the assumptions inherent in a partial derivative. For example,
  results will only be as valid as the assumption that g$_1$ and uWUE
  (and by extension $\lambda$) are fixed with respect to the user's
  conceptual change in VPD. Care must also be taken with possible
  indirect effects associated with a change in VPD: for example, a
  change in ET induced by a change in VPD can cause a change in
  surface temperature, which would drive a change in net
  radiation. These types of indirect effects and feedbacks are not
  considered in Eq. (\ref{d_et}): temperature (a variable) is
  mathematically fixed.

@@

The ecosystem-scale results of \citeA{Medlyn_2017} did not display the
robust relationship between plant type and and plant parameter (g$_1$)
that was demonstrated by \citeA{Zhou_2015} for uWUE. Given that there
is a robust relationship between g$_1$ and plant type in the
leaf-scale results of \citeA{Lin_2015}, and the ecosystem-scale
results of \citeA{Zhou_2015} rely on the same underlying theory as
\citeA{Medlyn_2017}, the ambiguous results in \citeA{Medlyn_2017}
could be due to noise in ecosystem scale observations, the
consequences of imposing a model structure in some of their g$_1$
estimations, or real ecosystem-scale within-plant type variability
(e.g. multiple species impacting the relationship within the
eddy-covariance footprint). These differences between leaf-scale
behavior and ecosystem scale behavior highlight the importance of
understanding how and why slope parameters show increased variability
at the ecosystem scale, as argued by \citeA{Medlyn_2017}. However, if
we assume that there is some real analogy between leaf scale results
in \citeA{Lin_2015} and ecosystem scale behavior, and that some of the
ambiguity in \citeA{Medlyn_2017} is due to model and/or observational
error, then we can use the relationships between g$_1$ and plant
characteristics defined by \citeA{Lin_2015} to frame the ET response
to VPD. This is subject to the strong caveat that we still do not
fully understand g$_1$ behavior at the ecosystem scale. Extrapolating
the results of \citeA{Lin_2015} (e.g. Fig. 2) to the ecosystem scale
would reveal the following relationships between g$_1$ and plant
types:

@@

So far we have discussed the role of different parameters in
modulating the ET response. However, the form of the ET response is
imposed by our choice of stomatal conductance model. We now examine
how stomatal model choice, and specifically the exponent of the VPD
dependence in the stomatal conductance model, can impact the general
form of the ET-VPD relationship. There is a theoretical basis for the
square root VPD dependence in both the stomatal conductance model and
uWUE based on the assumption that stomata behave to maximize carbon
gain while minimizing water loss, which observations also generally
support \cite{Lloyd_1991, MEDLYN_2011, Lin_2015, Zhou_2014, Zhou_2015,
  Medlyn_2017}. However, some purely empirical results that fit the
exponent of the VPD dependence to data have shown that it may vary
slightly from 1/2, suggesting that stomata, as well as ecosystem-scale
quantities based on stomata theory, may not always function optimally
\cite{Zhou_2015, Lin_2018}. Specifically with regards to uWUE, one
would not expect that this ecosystem scale WUE quantity will respond
to VPD exactly analogously to stomata. Direct soil evaporation's
contributions to ET should shift the exponent of the VPD
dependence. The results from \citeA{Zhou_2015} corroborate this: they
found a mean empirically fit exponential VPD dependence of 0.55,
varying slightly from the theoretically optimal value of of 0.5 for
AmeriFlux sites. Results in \citeA{Lin_2018} also show variance in the
empirical exponent of the VPD dependence in the stomatal conductance
model, but interpretation of this variance is more difficult as
\citeA{Lin_2018} do not handle the GPP dependence of stomatal
conductance in a directly analogous manner to the optimal theory in
\citeA{MEDLYN_2011} and \citeA{Medlyn_2017}. Regardless, given that
these recent results on the relationship between VPD, GPP, and ET
\cite{MEDLYN_2011, Zhou_2014, Zhou_2015, Medlyn_2017} form the
backbone of our analysis and are what allowed us to derive an explicit
ET expression for the first time (Eq. (\ref{et})), we will analyze if
and how assumptions about the exponent of the VPD dependence impact
the shape of the ET-VPD dependence. This analysis is also important to
understand whether the choice of stomatal conductance model alters the
fundamental behavior of the ET-VPD relationships, as many commonly
used models utilize a VPD exponent other than the 1/2 suggested by
optimal theory \cite<e.g. >[ which uses an exponent of
1]{Leuning_1990}. It is useful to understand how the choice of model
imposes the functional form of the ET-VPD relationship, given its
fundamental importance for land-atmosphere interactions.

@@

\begin{figure}
  \centering
  \centerline{\includegraphics[width=0.75\textwidth]{./concave.pdf}}
  \caption{ Solutions corresponding to inflection points between
    concave up and concave down ET-VPD curves (Eq. (\ref{curves})) for
    three specific scenarios. Solutions are defined in terms of a
    non-dimensional VPD ($VPD^m/g_*$), but to aide physical
    interpretation the horizontal axis is additionally provided in
    terms of dimensionalized VPD assuming $m=1/2$ and
    $g_*=110\; Pa^{1/2}$ (average of all PFT $g_1$). The vertical axis
    has a different interpretation depending on the solution
    curve. For the blue line ($m$ varying), it corresponds to $m$, for
    the orange line ($n$ varying) it corresponds to $n$, and for the
    green line it corresponds to the value of both $n$ and $m$
    ($n=m$). Regions of the parameter space that correspond to concave
    up and concave down results are labeled: for each curve, the
    parameter space ``below'' the line corresponds to a concave up
    relationship, which the parameter space ``above'' the line
    corresponds to a concave down relationship.}
  \label{concave}
\end{figure}

@@

While it is possible (and even likely) that in the future some other
form of VPD dependence is derived, at present \cite{MEDLYN_2011} and
\cite{Zhou_2014} established n=m=1/2 as the most likely candidate
given current theory and empirical data. Additionally, we argue that a
concave up result matches physical intuition more than a concave down
result. Plants must maintain nutrient and sugar transport through the
phloem and xylem. To accomplish this, stomata must remain slightly
open \cite{De_2013, Nikinmaa_2013, Ryan_2014}. Furthermore, even if
complete stomatal closure were possible, cuticular water loss and (at
the ecosystem-scale) direct soil evaporation are still sources of ET
which increase with VPD, independent of stomatal closure. Therefore,
in the limit as VPD becomes large and we assume plants are exercising
all strategies to reduce ET, any further increase in VPD should result
in an increase in ET through cuticular water loss and/or direct soil
evaporation. This inevitable transition from conditions when stomata
respond strongly to VPD to conditions when stomata response is
asymptoting towards full closure would cause a concave up ET-VPD
curve, which is matched by our theory. In short, plant response
becomes more limited as VPD increases, while atmospheric demand
monotonically increases with VPD, leading to the result that
atmospheric demand dominates plant response when atmospheric demand is
high.

This analysis allows us to understand the theoretical shape of the ET
response to VPD with environmental conditions held
fixed. Accomplishing this with purely statistical methods applied to
flux observations would be very difficult, given the relatively fast
time scale of plant response and the non-stationarity of (solar
forced) environmental conditions over the relatively coarse (half
hourly) flux estimates (which is required to obtain robust
eddy-covariance statistics). Our results on the shape of the ET-VPD
curve with environmental conditions held fixed can be built upon with
future work examining how changes in VPD and environmental conditions
(e.g. soil water storage) feedback upon one another. In the soil water
storage example, over very long time scales extremely high VPD
perturbations coupled with no precipitation could result in decreases
in soil water storage such that water becomes limiting. This could be
represented by an extension of our framework in which $\lambda$ is
allowed to decrease with decreasing soil water, increasing uWUE and
decreasing g$_1$. Here, we focus our results by framing them with
previous literature to build baseline intuition for ET-VPD dependence.

@@

\acknowledgments This work used eddy covariance data acquired and
shared by the FLUXNET community, including these networks: AmeriFlux,
AfriFlux, AsiaFlux, CarboAfrica, CarboEuropeIP, CarboItaly, CarboMont,
ChinaFlux, Fluxnet-Canada, GreenGrass, ICOS, KoFlux, LBA, NECC,
OzFlux-TERN, TCOS-Siberia, and USCCC. The ERA-Interim reanalysis data
are provided by ECMWF and processed by LSCE. The FLUXNET eddy
covariance data processing and harmonization was carried out by the
European Fluxes Database Cluster, AmeriFlux Management Project, and
Fluxdata project of FLUXNET, with the support of CDIAC and ICOS
Ecosystem Thematic Center, and the OzFlux, ChinaFlux and AsiaFlux
offices. We would like to thank Dr. Trevor Keenan for providing tools
for citation of the FLUXNET2015 dataset, available at
\url{https://github.com/trevorkeenan/FLUXNET_citations}. This material
is based upon work supported by the National Science Foundation
Graduate Research Fellowship under Grant No. DGE 16-44869. We would
like to thank Emma Robinson and one anonymous reviewer for thoughtful
comments that greatly improved the strength of the manuscript.

### 2019-09-04 doc/paper/james-supplement.tex
done with response to review 1

@@

>>>>

Additionally, we include a figure showing the joint distribution
between saturation vapor pressure and relative humidity calculated
from the FLUXNET-2015 data. Relative humidity and saturation vapor
pressure are much more independent than saturation vapor pressure and
VPD, and we use an assumption that relative humidity and saturation
vapor pressure can be approximated as independent in order to evaluate
$\frac{\partial ET}{\partial VPD}$. Please note that at a given site,
the relationship may be more or less independent depending on the
hydroclimate.

@@

\begin{figure}
  \centering \includegraphics{./supp-figs/0joint_rh_es.pdf}
  \caption{The joint distribution of relative humidity and saturation
    vapor pressure for the FLUXNET2015 dataset.}
  \end{figure}
